## Pre-Workshop Exercise

Please write the code for the following pipeline steps:

1.  Load the `tidyverse` and `readxl` libraries.
2.  Read in the Excel file located in: `"data/medicaldata_tumorgrowth.xlsx"` into a variable `tumor`.
3.  Select the all the columns except `Grp`, and filter the rows such that Day is `0` or `20`. Save this data subset into a variable `tumor_subset`.
4.  We want to compare baseline tumor sizes (Day 0) with tumor sizes at Day 20 between each of the groups.
    - Using `tumor_subset`, calculate the average tumor `Size` for each `Group` and `Day`.
5.  Save `tumor_subset` into a CSV file located in `"data/tumorsubset.csv"`.

```{r pre-workshop, echo=FALSE, exercise=TRUE, exercise.lines = 20}

```

```{r, include=FALSE}
# solution
library(tidyverse)
library(readxl)

.tumor <- read_excel("data/medicaldata_tumorgrowth.xlsx")

.tumor_subset <- .tumor %>%
  select(-Grp) %>%
  filter(Day == 0 | Day == 20)

.tumor_subset %>%
  group_by(Group, Day) %>%
  summarize(mean(Size))

write_csv(.tumor_subset, "data/tumorsubset_solution.csv")
```
